---
title: "Assignment 3"
output: pdf_document
---

## Problem 1
Obtain 100 exponential random variables (with rate 1) and arrange them by column 
in a matrix called M with 10 rows and 10 columns. Use seed 3100.
```{r}
set.seed(3100)  
M <- matrix(rexp(100, rate = 1), nrow = 10, ncol = 10) 
M
```

a. Find the transpose of M.
```{r}
M_transpose <- t(M)  
M_transpose
```

b. Find (M transpose) x M using matrix multiplication.
```{r}
M_trans_M <- M_transpose %*% M
M_trans_M
```

c. Find 2.5M + 5.2I10, where I10 is an identity matrix that has 10 rows and 10 columns.

```{r}
I_10 <- diag(10) 
sum <- (2.5 * M) + (5.2 * I_10)  
sum
```

d. Find the inverse matrix of M.
```{r}
# Let's first check if M has an inverse by calculating its determinant
det_M <- det(M) 
det_M # As long as det_M is non-zero, it will have an inverse

M_inverse <- solve(M)
M_inverse
```

e. Find the overall mean of the elements in M.
```{r}
overall_mean <- mean(M)  
overall_mean
```

f. Using which.max(), find the row of M with the largest sum.
```{r}
row_sums <- apply(M, 1, sum)
row_w_largest_sum <- which.max(row_sums)
row_w_largest_sum
```

g. For each column in M, find the percentage of elements greater than 1.
```{r}
percentage_above_1 <- apply(M, 2, function(x) sum(x > 1) / length(x) * 100)
percentage_above_1 # Since there are 10 entries per column, expect to see a multiple of 10 
```

h. Using string functions, give the names “n1”, “n2”, …, “n10” to the rows of M, 
and “v1”, “v2”, …, “v10” to the columns of M. Do not type out the name vectors.
```{r}
rownames(M) <- paste0("n", 1:10)
colnames(M) <- paste0("v", 1:10)
rownames(M)
colnames(M)
```

i. Extract a sub matrix of M consisting of rows “n3”, “n4”, and “n5” and all 
columns except “v2” and “v4”.
```{r}
sub_matrix <- M[c("n3", "n4", "n5"), c(1, 3, 5:10)]
sub_matrix
```

j. Create a diagonal matrix L that has the same diagonal elements as M.
```{r}
L <- diag(diag(M))
L
```

k. Add to L an additional row of all zeros after the last row.
```{r}
L_add_row <- rbind(L, rep(0, ncol(L))) # ncol() outputs the no. of columns
L_add_row
```

l. Extract the last column of M as a 1-column matrix and the last row of M as a 1-row matrix. 
Form a new matrix H by multiplying the last column of M by the last row using matrix multiplication.
```{r}
last_column <- M[, ncol(M), drop = FALSE]  
last_row <- M[nrow(M), , drop = FALSE] 
last_column
last_row

H <- last_column %*% last_row
H
```
