---
title: "Assignment 8"
output: pdf_document
---

The following code is used to prepare the dataset:
```{r}
library(tidyverse)
library(lubridate)

circ <- read_csv("Charm_City_Circulator_Ridership.csv")

# Convert dates
circ <- circ |> mutate(date = mdy(date))

# Change column names for reshaping
colnames(circ) <- colnames(circ) |>
  str_replace("Board", ".Board") |>
  str_replace("Alight", ".Alight") |>
  str_replace("Average", ".Average")

# Make long format by route/type
long <- circ |>
  pivot_longer(
    c(
      starts_with("orange"),
      starts_with("purple"),
      starts_with("green"),
      starts_with("banner")
    ),
    names_to = "var",
    values_to = "number"
  )

# Separate route/type into route and type
long <- long |>
  separate(var, into = c("route", "type"), sep = "[.]")

# Make wide format by type
type_wide <- long |>
  spread(type, value = number)

# Extract just average ridership per day
avg <- long |>
  filter(type == "Average", !is.na(number)) |>
  select(-type) |>
  rename(Average = number)
```

# Problem 1
Plot average ridership by date using a scatterplot
```{r}
ggplot(avg, aes(x = date, y = Average, color = route)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Average Ridership by Date",
    x = "Date",
    y = "Average Ridership",
    color = "Route") 

ggplot(avg, aes(x = date, y = Average, color = route)) +
  geom_point(alpha = 0.5) +  
  geom_smooth(se = FALSE, color = "black", aes(group = route)) +
  labs(
    title = "Average Ridership by Date",
    x = "Date",
    y = "Average Ridership",
    color = "Route")
```

# Problem 2
Color the points by day of the week.
```{r}
ggplot(avg, aes(x = date, y = Average, color = day)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Average Ridership by Date",
    x = "Date",
    y = "Average Ridership",
    color = "Day of the Week")
```

# Problem 3
Color the points by route, using the following palette:
pal <- c("darkorange", "purple", "darkgreen", "blue")
```{r}
pal <- c("darkorange", "purple", "darkgreen", "blue")

ggplot(avg, aes(x = date, y = Average, color = route)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = pal) +
  labs(
    title = "Average Ridership by Date",
    x = "Date",
    y = "Average Ridership",
    color = "Route") 
```

# Problem 4
Plot average ridership by date in separate panels by route.
```{r}
ggplot(avg, aes(x = date, y = Average)) +
  geom_point(alpha = 0.5) + 
  facet_wrap(~ route) +
  labs(
    title = "Average Ridership by Date",
    x = "Date",
    y = "Average Ridership")
```

# Problem 5
Plot average ridership by date in separate panels by day of the week. Color the points by route.
```{r}
ggplot(avg, aes(x = date, y = Average, color = route)) +
  geom_point(alpha = 0.5) +  
  facet_wrap(~ day) + 
  labs(
    title = "Average Ridership by Date",
    x = "Date",
    y = "Average Ridership",
    color = "Route") 
```

# Problem 6
Plot average ridership by date. Color the points by route. Customize the plot:
Change the x-axis label to “Year”.
Change the y-axis label to “Number of People”.
Use the black-and-white theme: theme_bw().
Set the text size using theme(text = element_text(size = 20)).
```{r}
ggplot(avg, aes(x = date, y = Average, color = route)) +
  geom_point(alpha = 0.5) + 
  labs(
    title = "Average Ridership by Date",
    x = "Year",  
    y = "Number of People",  
    color = "Route") +
  theme_bw() +  
  theme(text = element_text(size = 20))
```

# Problem 7
Plot average ridership on the orange route vs. date as a solid line. 
Add dashed lines for Boardings and Alightings.
Use “darkorange” for all line colors.
Use linetype as an aesthetic for lines.
Customize line types using scale_linetype_manual() with:
values = c(Average = "solid", Boardings = "dashed", Alightings = "dashed")

```{r}
ggplot(filter(long, route == "orange"), aes(x = date, y = number, linetype = type)) + 
  geom_line(color = "darkorange", size = 0.3) + 
  scale_linetype_manual(values = c("Average" = "solid", "Boardings" = "dashed", "Alightings" = "dashed")) +
  labs(
    title = "Ridership by Date (Orange)",
    x = "Date",
    y = "Number of People",
    linetype = "Line Type")

# Assess whether the boardings and alightings differ from the average values.
```
The dashed lines for Boardings and Alightings are hard to see because 
the solid line overrides them. So, this means there isn't a big difference
between the boardings and alightings from the average values. 




