---
title: "Assignment 1"
output: pdf_document
---

## Problem 1
Calculate the remainder after dividing 31079 into 170166719.
```{r}
# use the modulo operator %% to calculate the remainder
remainder <- 170166719 %% 31079
remainder
```

## Problem 2
Calculate the respective areas of circles having radii 3, 4, …, 100.
```{r}
# area of circle is radius sqaured times pi
areas <- pi * (3:100)^2
# output will have 97 values since the sequence 3:100 consists of 97 values
areas
```

## Problem 3
Calculate the interest earned on an investment of $2000, assuming an interest rate of 3% compounded annually, for terms of 1, 2, …, 30 years.
```{r}
net_worth <- 2000 * (1.03^(1:30))
# but what we're really interested is the "interest earned", so the profit
interest <- net_worth - 2000
interest
```

## Problem 4 
Using the rep() and seq() functions as needed, create the following vectors:
{ 0 0 0 0 0 1 1 1 1 1 2 2 2 2 2 3 3 3 3 3 4 4 4 4 4 }
{ 1 2 3 4 5 1 2 3 4 5 1 2 3 4 5 1 2 3 4 5 1 2 3 4 5 }
{ 1 2 3 4 5 2 3 4 5 6 3 4 5 6 7 4 5 6 7 8 5 6 7 8 9 }

```{r}
# the default setting of the rep() function is as follows: rep(x, times = 1, length.out = NA, each = 1)
# for this problem we just need to manipulate the "times" and "each" argument
x <- rep(0:4, each = 5)
x

y <- rep(1:5, times = 5)
y

z <- rep(1:5, times = 5) + rep(0:4, each = 5) # no need to worry about recycling since total # of values match
z
```

## Problem 5
Find the value of r^1 + r^2 + ... + r^n, for all values of n between 1 and 100, when r = 1.08.
```{r}
r <- 1.08 # the default r value set by the problem

exponents <- 1:100 # this is a sequence of values from 1 to 100

# sum of a geometric series uses the formula S = a * (1 - r^n) / (1 - r)
sum_values <- (r * (1 - r^exponents)) / (1 - r)
sum_values
```

## Problem 6
Identify the elements of the sequence {2^1, …, 2^15} that exceed the corresponding elements of the sequence {1^3, …, 15^3}.
```{r}
# Let's first create the sequences and then compare them
seq1 <- 2^(1:15)  
seq2 <- (1:15)^3

result <- seq1 > seq2 # this inequality will spit out a boolean value
# if seq1 > seq2 is true, it will output "true"
result
# so the 1st & 10~15th element of the first seq. exceeds the corresponding elements of the second seq
```

## Problem 7
The following are a sample of observations on incoming solar radiation at a greenhouse:
{11.1, 10.6, 6.3, 8.8, 10.7, 11.2, 8.9, 12.2 }

### Assign the data to an object called solar.radiation.
``` {r}
solar.radiation <- c(11.1, 10.6, 6.3, 8.8, 10.7, 11.2, 8.9, 12.2)
``` 
### Find the mean, median, range, and variance of the radiation observations.
```{r}
mean_solar <- mean(solar.radiation)
median_solar <- median(solar.radiation)
range_solar <- range(solar.radiation)
variance_solar <- var(solar.radiation)
mean_solar
median_solar
range_solar
variance_solar
```
### Add 10 to each observation of solar.radiation and assign the result to sr10. Find the mean, median, range, and variance of sr10. Which statistics change, and by how much?
```{r}
sr10 <- solar.radiation + 10

mean_sr10 <- mean(sr10)
median_sr10 <- median(sr10)
range_sr10 <- range(sr10)
variance_sr10 <- var(sr10)
mean_sr10 
median_sr10 
range_sr10 
variance_sr10 

# Adding 10 increases the mean, median, and range (both lower & upper limit) by that amount
# Variance is unchanged
```
### Multiply each observation of solar.radiation by -2 and assign the result to srm2. Find the mean, median, range, and variance of srm2. How do the statistics change now?
```{r} 
srm2 <- solar.radiation * -2

mean_srm2 <- mean(srm2)
median_srm2 <- median(srm2)
range_srm2 <- range(srm2)
variance_srm2 <- var(srm2)
mean_srm2
median_srm2
range_srm2
variance_srm2

# Multiplying by -2 multiplies the mean, median by that amount
# Similar thing happens for range, but since the sign is changed, lower & upper limits swap places
# Variance multiplies by the square of the (-2), that is 4
```