---
title: "Assignment 4"
output: pdf_document
---

## Problem 1
Create a list named movie_info to store the following information about a movie:
  Title: "Inception"
  Release year: 2010
  Box office revenue (in millions): 829.9
  Rotten Tomatoes scores: Tomatometer = 87, Audience = 91
  Cast: "Leonardo DiCaprio", "Joseph Gordon-Levitt", "Ellen Page", "Tom Hardy"
```{r}
movie_info <- list(
  Title = "Inception",
  Release_Year = 2010,
  Box_Office_Revenue = 829.9,  
  Rotten_Tomatoes_Scores = list(Tomatometer = 87, Audience = 91),
  Cast = c("Leonardo DiCaprio", "Joseph Gordon-Levitt", "Ellen Page", "Tom Hardy")
)
```

a. Use str() to inspect the structure of the list.
```{r}
str(movie_info)
```

b. Add a new element to the list: the director's name ("Christopher Nolan").
```{r}
movie_info$Director <- "Christopher Nolan"
movie_info
```

c. Access only the Tomatometer score using both $ and [[ indexing.
```{r}
# Need to index twice since Tomatometer itself is a list as well
movie_info$Rotten_Tomatoes_Scores$Tomatometer
movie_info[["Rotten_Tomatoes_Scores"]][["Tomatometer"]]
movie_info
```

d. Create another list for another movie of your choice and combine the two lists 
into a single list of movies named movie_collection.
```{r}
movie_info2 <- list(
  Title = "The Conjuring",
  Release_Year = 2013,
  Box_Office_Revenue = 319.5,
  Rotten_Tomatoes_Scores = list(Tomatometer = 86, Audience = 83),
  Cast = c("Vera Farmiga", "Patrick Wilson", "Lili Taylor", "Ron Livingston"),
  Director = "James Wan"
)
movie_info2

# Create a nested list
movie_collection <- list(Inception = movie_info, Conjuring = movie_info2)
movie_collection
```

e. Use sapply() to extract the titles of all movies in movie_collection.
```{r}
movie_titles <- as.character(sapply(movie_collection, function(x) x$Title))
# without as.character, movie_titles will be a list
movie_titles
```
