---
title: "Assignment 5"
output: pdf_document
---

## Problem 1
Perform the following tasks related to dates and times in R:

a. Use Sys.Date() to get the current date and assign it to a variable named today.
```{r}
today <- Sys.Date()
today
```

b. Convert the following string to a Date object using as.Date(): "12/31/2023". 
Use the format "%m/%d/%Y".
```{r}
Date <- as.Date("12/31/2023", format = "%m/%d/%Y")
Date
```

c. Create a sequence of dates starting from January 1, 2024, to January 10, 2024, 
using the seq.Date() function with a daily interval.
```{r}
date_sequence <- seq.Date(from = as.Date("2024-01-01"), 
                          to = as.Date("2024-01-10"), by = "day")
date_sequence
```

d. Calculate the number of days between today and "1986-01-28".
```{r}
days_difference <- today - as.Date("1986-01-28")
days_difference
class(days_difference)
```

e. Create a POSIXct date-time object for "2023-12-25 15:30:00" in the UTC time zone.
```{r}
datetime_utc <- as.POSIXct("2023-12-25 15:30:00", tz = "UTC")
datetime_utc
```

f. Add 7 days and 12 hours to the date-time object created in part (e).
```{r}
new_datetime <- datetime_utc + (7 * 24 + 12) * 3600  
new_datetime
```

g. Using difftime(), calculate the difference between "2023-12-31 23:59:59" and 
"2023-01-01 00:00:00" in terms of weeks.
```{r}
start_time <- as.POSIXct("2023-01-01 00:00:00", tz = "UTC")
end_time <- as.POSIXct("2023-12-31 23:59:59", tz = "UTC")
time_diff <- difftime(end_time, start_time, units = "weeks")
time_diff
```

h. Convert the following times into a difftime object:
"3:45" as hours and minutes.
"2:30:15" as hours, minutes, and seconds.
```{r}
time_1 <- as.difftime("3:45", format = "%H:%M", units = "mins")
time_1
time_2 <- as.difftime("2:30:15", format = "%H:%M:%S", units = "secs")
time_2
```

i. Write an R script to determine if a given date (e.g., "2024-02-29") is valid 
and falls on a leap year.
```{r}
date_validity <- function(date_str) {
  date_input <- as.Date(date_str, format = "%Y-%m-%d")
  if (is.na(date_input)) {
    return(cat(date_str, "is NOT a valid date.\n"))
  }
  return(cat(date_str, "is a valid date.\n"))
}

# If the year is divisible by 4 and not divisible by 100, then it is a leap year
# If the year is divisible by both 4 and 100, then it should also be divisible by 
# 400 to be considered a leap year

is_leap_year <- function(date_str) {
  year <- as.numeric(format(as.Date(date_str), "%Y"))
  return ((year %% 4 == 0) & (year %% 100 != 0)) | (year %% 400 == 0)
}
```

j. Using the current system time, extract and print the year, month, and day 
components individually.
```{r}
year <- format(today, "%Y")
month <- format(today, "%m")
day <- format(today, "%d")

cat("Year:", year, "Month:", month, "Day:", day)
```

## Problem 2 
USArrests

a. Determine the number of rows and columns.
```{r}
nrow(USArrests)
ncol(USArrests)
```

b. Show the row corresponding to Florida. Access the row by name, not number.
```{r}
FL_row <- USArrests["Florida", ]
FL_row
```

c. Show the last three rows.
```{r}
last_three_rows <- tail(USArrests, 3)
last_three_rows
```

d. Find the name of the states in rows 20, 30, and 40.
```{r}
state_names <- rownames(USArrests)[c(20, 30, 40)]
state_names
```

e. Create two new data frames (keeping the same row names as USArrests):
USArrests2 containing columns Murder, Assault, and Rape.
Urban that only contains the column UrbanPop.
```{r}
USArrests2 <- USArrests[, c("Murder", "Assault", "Rape")]
USArrests2
Urban <- USArrests[, "UrbanPop", drop = FALSE]
Urban
```

f. In Urban, rename the column UrbanPop to “UrbanPercent”.
```{r}
colnames(Urban)[colnames(Urban) == "UrbanPop"] <- "UrbanPercent"
colnames(Urban)
```

g. Find the median of each column in USArrests2.
```{r}
column_medians <- apply(USArrests2, 2, median)
column_medians
```

h. In USArrests2, add a column called “Total”, which gives the total sum of Murder, 
Assault, and Rape.
```{r}
USArrests2$Total <- rowSums(USArrests2)
USArrests2
```

i. Using the order() function, list the states in USArrests2 according to Murder
in decreasing order.
```{r}
ordered_states <- rownames(USArrests2)[order(USArrests2$Murder, decreasing = TRUE)]
ordered_states
```

j. Show the subset of the data frame USArrests2 containing all states having 
Murder < 10, Assault < 100, and Rape < 10.
```{r}
subset_USArrests2 <- USArrests2[USArrests2$Murder < 10 & USArrests2$Assault < 100 & USArrests2$Rape < 10, ]
subset_USArrests2
```

k. Using data frame USArrests, find the across-state average murder rate (Murder) 
in regions where the percentage of the population living in urban areas (UrbanPop) 
exceeds 77%. Compare this with the average murder rate where the urban area 
population is less than 50%.
```{r}
high_urban_avg_murder <- mean(USArrests$Murder[USArrests$UrbanPop > 77])
low_urban_avg_murder <- mean(USArrests$Murder[USArrests$UrbanPop < 50])
cat("Average Murder Rate (UrbanPop > 77%):", high_urban_avg_murder, "\n")
cat("Average Murder Rate (UrbanPop < 50%):", low_urban_avg_murder, "\n")
```

l. Use the plot() function to show Murder vs Assault. 
Identify and explain any relationships you see.
```{r}
plot(USArrests$Murder, USArrests$Assault, 
     main = "Murder vs Assault (per 100,000 people)", 
     xlab = "Murder Rate", 
     ylab = "Assault Rate", 
     pch = 19, col = "blue")

# States with higher rates of murder also tend to have higher rates of assault.
# More states fall toward the lower end of the murder & assault axes.
```

# Problem 3
Install the package COUNT. Then run the following:
```{r}
library(COUNT)
data(fishing)
pois.glm <- glm(totabund ~ meandepth, data = fishing, family = poisson)
```

a. Verify that the object pois.glm is a list.
```{r}
typeof(pois.glm)
```

b. Find the number of components in pois.glm and display their names.
```{r}
length(names(pois.glm))
names(pois.glm)
```

c. Among all the components of pois.glm, is there any component which is a list itself?
```{r}
typeof(pois.glm$qr) 
# qr is an example of a component which is a list itself
```

d. Find out the number of elements contained in the component residuals, 
as well as the maximum and minimum values.
```{r}
pois.glm$residuals
length(pois.glm$residuals)
max(pois.glm$residuals)
min(pois.glm$residuals)
```

e. Add a new component named “extra” to pois.glm, containing 10 random numbers 
following a Poisson(2) distribution.
```{r}
new_component <- rpois(10, lambda = 2)
pois.glm$extra <- new_component
```



