---
title: "Employee Mental Health & Job Satisfaction"
output: pdf_document
---

Select the following packages 
```{r}
library(readxl)
library(ggplot2)
library(tidyverse)
```

# Part 1 
##. A. Mean, median, standard deviation, quantiles, min/max of numerical variables
```{r}
data <- read_excel("Group9_EmployeeMentalHealth_JobSatisfaction (1).xlsx")
selected_columns <- data[, c("Age", "Work_Experience", "Weekly_Work_Hours", "Stress_Level", "Work_Life_Balance", "Company_Culture_Score", "Job_Satisfaction" )]

# Compute summary statistics
summary_stats <- data.frame(
  Mean = sapply(selected_columns, mean),
  Median = sapply(selected_columns, median),
  Std_Dev = sapply(selected_columns, sd),
  Min = sapply(selected_columns, min),
  Q1 = sapply(selected_columns, function(x) quantile(x, 0.25)),
  Q3 = sapply(selected_columns, function(x) quantile(x, 0.75)),
  Max = sapply(selected_columns, max)
)

summary_stats
```

## B. Create boxplots for each numerical variable. 
```{r}
ggplot(data, aes(y = Age)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for Age", y = "Ages") 

ggplot(data, aes(y = Work_Experience)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for Work Experience", y = "Work Experience (years)") 

ggplot(data, aes(y = Weekly_Work_Hours)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for Weekly Work Hours", y = "Weekly Work Hours") 

ggplot(data, aes(y = Stress_Level)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for Stress Level", y = "Stress Level") 

ggplot(data, aes(y = Work_Life_Balance)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for Work Life Balance", y = "Work Life Balance") 

ggplot(data, aes(y = Company_Culture_Score)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for Company Culture Score", y = "Company Culture Score") 

ggplot(data, aes(y = Job_Satisfaction)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  labs(title = "Boxplot for Job Satisfaction", y = "Job Satisfaction") 
```
Outliers are colored in red.

## C. Normality Testing

First, let's inspect visually
```{r}
plot_normal_curve <- function(data, col_name) {
  ggplot(data, aes_string(x = col_name)) +
    geom_histogram(aes(y = after_stat(density)), bins = 30) +
    labs(title = paste("Histogram and Normal Curve for", col_name), 
         x = col_name, y = "Density") 
}

for (col in names(selected_columns)){print(plot_normal_curve(data, col))}
```
Upon inspection, only the histogram for Weekly Work Hours looks normally distributed.

Now, let's test normality using the Shaprio-Wilk test.
The null hypothesis for the Shapiro-Wilk test is that data "is" normally distributed.
```{r}
shapiro_test_results <- lapply(selected_columns, shapiro.test)
shapiro_test_results
```
All of the variables have a very small p-value, which means we have to reject
the null hypothesis and accept the alternative hypothesis, which is that 
data is not normally distributed. 

## D. Pearson’s correlation analysis to determine relationships between numerical variables.

```{r}
correlation_matrix <- cor(selected_columns, method = "pearson")
correlation_matrix
```

## E. Consider transforming or normalizing the data. 

Let's try z-score normalization
```{r}
z_scores <- sapply(selected_columns, function(x) {
  (x - mean(x)) / sd(x)
}, simplify = FALSE)

trans_shapiro_test_results <- lapply(z_scores, shapiro.test)
trans_shapiro_test_results
```
Even after z-score normalization, we can see how all the p-values are very small.
This means that we have to reject the null hypothesis again, so once again,
according to the Shapiro test, our transformed data is not normally distributed. 

# Part 2
Our goal now is to test whether employees in high-stress industries report lower 
job satisfaction scores than those in low-stress industries.

Since we failed to transform our data so that it follows normal distribution,
we will use the Mann-Whitney U test, which does not have that as a requirement.

Null Hypothesis (H0): There is no relationship between stress level and job satisfaction.
Alternative Hypothesis (H1): There is a relationship between stress level and job satisfaction.

```{r}
# Perform the Mann-Whitney U test (also known as Wilcoxon rank-sum test)
wilcox_test_results <- wilcox.test(data$Job_Satisfaction, data$Stress_Level)
wilcox_test_results
```
Based on the results, we can conclude that there is a statistically significant difference 
in job satisfaction scores between individuals with high and low stress levels. 
The p-value 2.2e-16 is extremely small, so we accept the alternative hypothesis.
2.2e-16 is much, much smaller than alpha=0.05, so we our findings are definetely
statistically significant at alpha=0.05

For the last step, let's get the correlation coefficient
```{r}
correlation_coefficient <- cor.test(data$Job_Satisfaction, data$Stress_Level, method = "pearson")
correlation_coefficient
```
Again, p-value = 5.057e-15 is extremely small so these are statistically significant
results. The correlation coefficient is -0.34, meaning that for every unit increase
in job stress level, job satisfaction decreases by 0.34. 

Hence, to answer the question, yes, employees who report high-stress lower satisfaction
scores. 


